---
title: "Individual 2"
author: "Andrew Sundberg"
date: "March 22, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Lab 4

## Part 1

```{r}
library(readxl)
defense = read_excel('C:/Users/Andrew Sundberg/Documents/DS202/cyclonesFootball2018.xlsx', sheet='Defensive')
offense = read_excel('C:/Users/Andrew Sundberg/Documents/DS202/cyclonesFootball2018.xlsx', sheet='Offensive')
biography = read_excel('C:/Users/Andrew Sundberg/Documents/DS202/cyclonesFootball2018.xlsx', sheet='Biography')
```

##1. Factorize variables

```{r}
defense$Name = factor(defense$Name)
defense$Opponent_Opponent = factor(defense$Opponent_Opponent)
offense$Name = factor(offense$Name)
offense$Opponent_Opponent = factor(offense$Opponent_Opponent)
biography$Name = factor(biography$Name)
```

##2. Make weight numeric
```{r}
biography$Weight = as.numeric(biography$Weight)

library(dplyr)
defClean = defense %>% mutate_at(c(3:11), as.numeric)
offClean = offense %>% mutate_at(c(3:12), as.numeric)
biography = biography %>% mutate_at(c(4), as.numeric)

biography$Height = sapply(strsplit(as.character(biography$Height),"-"),
        function(x){12*as.numeric(x[1]) + as.numeric(x[2])}) #convert to inches
bioClean <- biography
```


## Part 2: Tidying

## 1. Reformat defClean
```{r}
library(tidyr)
defClean2 <-defClean %>%
  gather(key=Statistic, value=count, 3:11) 
```


##2. Histogram

```{r}
library(ggplot2)
ggplot(defClean2, aes(x=count)) +geom_histogram(bins = 10) + facet_wrap(~Statistic) 
```

Defensive stats that are obviously rare are forced fumbles, fumble recoveries, interceptions, pass break-ups and sacks.  These histograms do not have many players outside of the first row and are very close to zero.


##3. K State or West Virginia

```{r}
defense2 = defClean2 %>%
  filter(Opponent_Opponent %in% c('Kansas State', 'West Virginia') & Statistic == "Tackles_Solo") %>%
  select(Name, Opponent_Opponent, Statistic, count) %>%
  spread(key = Opponent_Opponent, value = count)

str(defense2)

names(defense2) = c("Name", "Statistic", "KState_Tackles", "WV_Tackles")

ggplot(defense2, aes(x = WV_Tackles, y = KState_Tackles)) + geom_point()
``` 


It looks like we played better defense against Kansas State.  The scale on the y-axis is much larger and if you add up all the tackles throughout the whole team, ther are more team tackles against Kansas State.  Therefore, we played better defense against Kansas State.



##4. Separate Hometown

```{r}
hometown = bioClean %>% select(Hometown)
hometown = hometown %>% separate(Hometown, into=c('City', 'State'), sep=', ') 
```

##5. How many players from each state

```{r}
hometown %>% group_by(State) %>% summarize(Number_of_Players=n()) 
```


##Joining

##1. Join offense and bio

```{r}
off_bio = (left_join(offClean, bioClean,  by='Name'))
```

##2. Weight and Recieving yards
```{r}
library(ggplot2)
ggplot(off_bio, aes(x=Weight, y=Receiving_YDS)) +geom_point()
```

According to the graph produced, there appears to be no relationship between weight and recieving yards.  The players who weigh 225 pounds have the biggest variation in receiving yards as they have our top wide recievers it appears.


##3. Do heavier players get more tackles

```{r}
def_bio = (left_join(defClean, bioClean,  by='Name'))
ggplot(def_bio, aes(x=Weight, y=Tackles_Solo)) +geom_point()
```


It appears that actually lighter players record more tackles.  There are a lot more lighter players, but the top end of the players looks like it has a negative linear relationship.
