---
title: "individual1"
author: "Grant Ruedy"
date: "3/14/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readxl)
defense <- read_excel('cyclonesFootball2018.xlsx', sheet='Defensive')
str(defense)

offense <- read_excel('cyclonesFootball2018.xlsx', sheet='Offensive')
str(offense)

biography <- read_excel('cyclonesFootball2018.xlsx', sheet='Biography')
str(biography)
```

##Cleaning
1.
```{r}
defense$Name <- factor(defense$Name)
defense$Opponent_Opponent <- factor(defense$Opponent_Opponent)

offense$Name <- factor(offense$Name)
offense$Opponent_Opponent <- factor(offense$Opponent_Opponent)

biography$Name <- factor(biography$Name)

```

2. & 3.
```{r}
library(dplyr)
defClean <-defense %>% mutate_at(c(3:11), as.numeric)
offClean <- offense %>% mutate_at(c(3:12), as.numeric)
biography <- biography %>% mutate_at(c(4), as.numeric)

biography$Height <-sapply(strsplit(as.character(biography$Height),"-"),
        function(x){12*as.numeric(x[1]) + as.numeric(x[2])}) #convert to inches
bioClean <- biography

str(defClean)
str(offClean)
str(bioClean)
```

##Tidying (1)

1. & 2.
```{r}
library(tidyr)
defClean <-defClean %>%
  gather(key=Statistic, value=count, 3:11) 
head(defClean)

library(ggplot2)
ggplot(defClean, aes(x=count)) +geom_histogram(bins = 10) + facet_wrap(~Statistic) 
```

Based off of the histograms for each defensive stat, interceptions, forced fumbles, and fumble recoveries (all for turnovers) are the most rare. These three statistics have almost every observation at zero, showing that few people achieve these during games. Sacks are also very rare, as only a small amount of players had 1 or 2 per game. Tackles seem to be the most common, which makes sense because they happen almost every play.

3.
```{r}
library(dplyr)

defClean_SoloTackles <- defClean[defClean$Statistic=='Tackles_Solo',]

a <-defClean_SoloTackles %>%
  group_by(Name)
 
  
complete <- filter(a, Opponent_Opponent=='West Virginia' | Opponent_Opponent=='Kansas State') 

complete1 <- complete %>% spread(key=Opponent_Opponent, value=count)
names(complete1) <- c('Name', 'Statisitic', 'KS_Tackles','WV_Tackles')

ggplot(complete1, aes(x=WV_Tackles, y=KS_Tackles), group_by(Name)) + geom_point()
```

Based on the graph, Iowa State had better defense against Kansas State. This is because it looks like each player had more tackles, because the highest amount of tackles was 7.5. The highest amount of tackles anyone had against West Virginia was only 5. Also, only two players against West Virginia had more than 2 tackles, while 5 players had more than 5 tackles againt Kansas State.

##Tidying (2)
1.
```{r}
hometown <- bioClean %>% select(Hometown)
hometown <- hometown %>% separate(Hometown, into=c('City', 'State'), sep=', ') 
head(hometown)
```

2.
```{r}
hometown %>% group_by(State) %>% summarize(Number_of_Players=n()) 
```


##Joining
1.
```{r}
joined <- (left_join(offClean, bioClean,  by='Name'))
joined

```

2.
```{r}
library(ggplot2)
ggplot(joined, aes(x=Weight, y=Receiving_YDS)) +geom_point()
```

Based on our graph, it doesn't seem like there is any noticable relationship between the weight of the player and how many receiving yards they get. Although players who weight 225 lbs seem to have the most receivng yards, implying that this is the best weight a receiving could be. Players who were the lightest (at 175 lbs) had the second most receving yards. This leads to the conclusion that if you are lighter (and presumably quicker), you have a good chace to have a lot of receiving yards.

3.
Question: Do taller players tend to score more receiving touchdowns?
```{r}
ggplot(joined, aes(x=Height, y=Receiving_TD, color=Position)) + geom_point()
```

Based on the graph, it looks like taller players score more touchdowns. The majority of players who scored one touchdown are over 75 inches (6 foot 3) tall. Also, the only player who scored 2 touchdowns in a game was 78 inches (6 foot 6) tall. This is very tall, even for a wide receiver, implying that this person used their height very much to their advantage by scoring mutliple touchdowns. It makes sense that the quarterback wants a big tall target to throw to in the end zone, to limit the chance of an interception by putting the ball where only the receiver can reach it. It is also interesting to point out that the two tight ends who scored a touchdown were 6 foot 3 and 6 foot 8. This implies it is important to have big tall tight ends, especially near the end zone.