---
title: "Team File"
author: "Grant Ruedy & Andrew Sundberg"
date: "3/25/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(readxl)
defense <- read_excel('cyclonesFootball2018.xlsx', sheet='Defensive')
str(defense)

offense <- read_excel('cyclonesFootball2018.xlsx', sheet='Offensive')
str(offense)

biography <- read_excel('cyclonesFootball2018.xlsx', sheet='Biography')
str(biography)
```

##Cleaning
1.
```{r}
defense$Name <- factor(defense$Name)
defense$Opponent_Opponent <- factor(defense$Opponent_Opponent)

offense$Name <- factor(offense$Name)
offense$Opponent_Opponent <- factor(offense$Opponent_Opponent)

biography$Name <- factor(biography$Name)

```


2. & 3.
```{r}
library(dplyr)
defClean1 <-defense %>% mutate_at(c(3:11), as.numeric)
offClean <- offense %>% mutate_at(c(3:12), as.numeric)
biography <- biography %>% mutate_at(c(4), as.numeric)

biography$Height <-sapply(strsplit(as.character(biography$Height),"-"),
        function(x){12*as.numeric(x[1]) + as.numeric(x[2])}) #convert to inches
bioClean <- biography

str(defClean1)
str(offClean)
str(bioClean)
```

##Tidying (1)

1. & 2.
```{r}
library(tidyr)
defClean <-defClean1 %>%
  gather(key=Statistic, value=count, 3:11) 
head(defClean)

library(ggplot2)
ggplot(defClean, aes(x=count)) +geom_histogram(bins = 10) + facet_wrap(~Statistic) 
```

Based off of the histograms for each defensive stat, interceptions, forced fumbles, and fumble recoveries (all for turnovers) are the most rare. These three statistics have almost every observation at zero, showing that few people achieve these during games. Sacks are also very rare, as only a small amount of players had 1 or 2 per game. Tackles seem to be the most common, which makes sense because they happen almost every play.

3.
```{r}
library(dplyr)

defClean_SoloTackles <- defClean[defClean$Statistic=='Tackles_Solo',]

a <-defClean_SoloTackles %>%
  group_by(Name)
 
  
complete <- filter(a, Opponent_Opponent=='West Virginia' | Opponent_Opponent=='Kansas State') 

complete1 <- complete %>% spread(key=Opponent_Opponent, value=count)
names(complete1) <- c('Name', 'Statisitic', 'KS_Tackles','WV_Tackles')

ggplot(complete1, aes(x=WV_Tackles, y=KS_Tackles), group_by(Name)) + geom_point()
```

Based on the graph, Iowa State had better defense against Kansas State. This is because it looks like each player had more tackles, because the highest amount of tackles was 7.5. The highest amount of tackles anyone had against West Virginia was only 5. Also, only two players against West Virginia had more than 2 tackles, while 5 players had more than 5 tackles againt Kansas State.

##Tidying (2)
1.
```{r}
hometown <- bioClean %>% select(Hometown)
hometown <- hometown %>% separate(Hometown, into=c('City', 'State'), sep=', ') 
head(hometown)
```

2.
```{r}
hometown %>% group_by(State) %>% summarize(Number_of_Players=n()) 
```

##Joining

##1. Join offense and bio

```{r}
off_bio = (left_join(offClean, bioClean,  by='Name'))
```

##2. Weight and Recieving yards
```{r}
library(ggplot2)
ggplot(off_bio, aes(x=Weight, y=Receiving_YDS)) +geom_point()
```

According to the graph produced, there appears to be no relationship between weight and recieving yards.  The players who weigh 225 pounds have the biggest variation in receiving yards as they have our top wide recievers it appears.


##3. Do heavier players get more tackles

```{r}
def_bio = (left_join(defClean1, bioClean,  by='Name'))
ggplot(def_bio, aes(x=Weight, y=Tackles_Solo)) +geom_point()
```


It appears that actually lighter players record more tackles.  There are a lot more lighter players, but the top end of the players looks like it has a negative linear relationship.
